Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-28 paper-ieee-trans.tex
Update paper-ieee-trans.tex

The source code was available at a private, centralized version control system.
The code was not validated by any Quality Assurance policy, but passed through a manual
certification procedure at the time of the release. This procedure tried to
improve the reliability of the software components by applying a set of
requirements [ref] to be fulfilled at each stage of the  development lifecycle (integration, certification,
pre--production, production), carried out by different teams. However, with 
EGEE--II, the developers started to use the the ETICS system [ref], solution aiming to develop
an automatic build system for Grid middleware.

@@

The E--Infrastructure for Testing, Integration and Configuration of Software
\cite{cordis:etics, cordis:etics2} (ETICS, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as EGEE and its gLite middleware. The framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto-generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The ETICS portal was the first automated
service for delivering quality software products in distributed environments like
the Grids.

@@

EGI--InSPIRE not only ensured the coordination of the operation of a secure, reliable 
European-wide production grid infrastructure federated from national grid initiatives 
that is integrated and interoperates with other grids worldwide but also foreseed the 
maintenance and development of operational tools, like the Operations Portal [ref], 
the EGI Helpdesk [ref] or Grid configuration repository (GOCDB) [ref] and the 
coordination of the of  the external provision of software required by EGI to 
form the production infrastructure. For the later it maintained a production--ready 
Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution) and 
defined general and component specific quality criteria to be applied to software 
components made available through this repository.
The role of {\sl UMD} was to enforce the fulfilment of this set of quality
criteria \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

The DevOps methods emphasize on exercising Quality Assurance techniques to
avoid infrastructure disruption whenever new developments are deployed into
production. The INDIGO--DataCloud project promoted the application of a 
Continuous Integration (CI) scenario to enforce the Quality Assurance requirements 
for any piece of software being produced within the project. Such environment 
requires an automation ready--to--go infrastructure where the different services involved
(source code management platform, automation server) interact with each other
to trigger the quality check execution and return back the exit codes, together
with their outputs. To accomplish this, the project leveraged on tightly
integrated open source tools such as GitHub [ref], as source code repository
and Jenkins [ref], as continuous integration system, as well as relying on
Docker container provisioning to speed up the source code validation. This CI
approach is guiding the project’s software development phase throughout the
first and second major releases.

@@

>>>>

The source code was available at a private, centralized version control system.
The code was not validated by any Quality Assurance policy, but passed through a manual
certification procedure at the time of the release. This procedure tried to
improve the reliability of the software components by applying a set of
requirements [ref] to be fulfilled at each stage of the  development lifecycle (integration, certification,
pre--production, production), carried out by different teams. However, with 
EGEE--II, the developers started to use the the ETICS system \cite{etics}, solution aiming to develop
an automatic build system for Grid middleware.

@@

The E--Infrastructure for Testing, Integration and Configuration of Software
\cite{cordis:etics, cordis:etics2} (ETICS, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as EGEE and its gLite middleware. The framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto-generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The ETICS framework was the first automated
service for delivering quality software products in distributed environments like
the Grids.

@@

EGI--InSPIRE not only ensured the coordination of the operation of a secure, reliable 
European-wide production grid infrastructure federated from national grid initiatives 
that is integrated and interoperates with other grids worldwide but also foreseed the 
maintenance and development of operational tools, like the Operations Portal \cite{egi-ops}, 
the EGI Helpdesk \cite{ggus} or Grid configuration repository (GOCDB) \cite{gocdb} and the 
coordination of the of  the external provision of software required by EGI to 
form the production infrastructure. For the later it maintained a production--ready 
Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution) and 
defined general and component specific quality criteria to be applied to software 
components made available through this repository.
The role of {\sl UMD} was to enforce the fulfilment of this set of quality
criteria \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

The DevOps methods emphasize on exercising Quality Assurance techniques to
avoid infrastructure disruption whenever new developments are deployed into
production. The INDIGO--DataCloud project promoted the application of a 
Continuous Integration (CI) scenario to enforce the Quality Assurance requirements 
for any piece of software being produced within the project. Such environment 
requires an automation ready--to--go infrastructure where the different services involved
(source code management platform, automation server) interact with each other
to trigger the quality check execution and return back the exit codes, together
with their outputs. To accomplish this, the project leveraged on tightly
integrated open source tools such as GitHub \cite{github}, as source code repository
and Jenkins \cite{jenkins}, as continuous integration system, as well as relying on
Docker container provisioning to speed up the source code validation. This CI
approach is guiding the project’s software development phase throughout the
first and second major releases.

@@

\bibitem{egi-ops}
\emph{EGI Operations Portal},
https://operations-portal.egi.eu/

\bibitem{ggus}
\emph{GGUS} EGI Helpdesk,
https://ggus.eu/

\bibitem{gocdb}
\emph{GOCDB} EGI Grid Configuration Repository,
https://ggus.eu/

\bibitem{github}
\emph{GITHUB},
https://github.com/

\bibitem{jenkins}
\emph{Jenkins} Continuous Integration System,
https://jenkins-ci.org/
